---
title: "Homework 1"
output: html_notebook
---

| Homework 1                     |               Data Programming in R |
| :---                           |                                ---: |
| Due by 11:59 PM on Tue Aug 30  | Business Analytics Graduate Program |
| Submit via GitHub              |     BAIS:6060 Fall 2022 (Part-Time) |

# Instructions

This homework covers to Module 1. Refer to the corresponding materials
and ICE (with solutions) posted on ICON.

To complete the homework, fill in the commands needed to finish all of
the exercises below. Program everything inside this notebook and do not
change the notebook's filename. If the exercises request that you store
information within certain variables, use those specific variables names
(case sensitive). This will assist in grading your homework in a timely
manner.

For notebooks in general:

- Comment your code so that a classmate, for example, could easily
follow along.

- Load any required packages using the `library(...)` command.
However, a notebook should *not* have any `install.packages(...)` or
`update.packages()` commands.

- Do not use the `View(...)` command in a notebook. Viewing a data
frame should be done interactively using RStudio's console but is not
recommended for a notebook that will be run by someone else.

For homeworks in general:

- Be careful not to alter any supplied data files. In particular, if you
open a CSV file with Excel, do *not* save it with Excel.

- Before turning in a homework, run your entire notebook by choosing
*Run > Restart R and Run All Chunks* in the RStudio interface.
Everything should work without error!

- To turn in a homework, unless otherwise instructed, upload *only* your
notebook to your corresponding GitHub repository. Please do not change
the notebook's filename. GitHub overwrites the previous version of the
file when you upload your version.

- If given a data set, there is no need to upload that data set back
into GitHub. Also, if your code generates any ouput files, such as an
HTML file, a new data set, or an image, then there is no need to upload
those files either.

Also included with this homework are the solutions within the included
file `soln.Rdata`. If you would like to see the solutions, execute
this command in the console: `load("soln.Rdata", soln <- new.env())`.
Then you can access the solutions within the object called `soln`.
For example, `View(soln$df)` will show the contents of the correct
data frame `df`, and `soln$fib` will show the results of the Fibonacci
exercise below. These are for your reference. The purpose of the
homework is for you to create these solutions on your own. Points will
be deducted if your turned-in notebook uses `soln.Rdata` in any way.

# Exercises

## #1

Save the value 5 in a variable called `KM` (for "kilometers"). Create
a new variable `MI` (for "miles"), which equals the value `0.62 * KM`.
Note that the variable names are all upper case and case sensitive.

```{r}
KM <- 5
MI <- 0.62 * KM

```

## #2

Use the exact command `mydata <- read.csv("data.csv")` to import
the included data file `data.csv` and store it in a variable called
`mydata`.

```{r}
mydata <- read.csv("data.csv")

```

## #3

Calculate the standard deviation of the data in column `A`, and store
the standard deviation in a variable called `stdev`.

```{r}
stdev <- sd(mydata$A)

```

## #4

Install the package called `numbers` (outside of this notebook), load it
(in this notebook), and use the package's `fibonacci(...)` function to
calculate the 23-rd Fibonacci number, i.e., plug in `n = 23` into the
`fibonacci(...)` function. Save the result in a variable called `fib`.

```{r}
library("numbers")
fib <- fibonacci(n=23)

```

## #5

Although pointing a laser at an aircraft is a federal crime, many people
still do it, and the FAA tracks all such reported incidents. Some of
these incidents are captured in the included file `lasers.csv`. Using
this file, complete the following:

### (a)

Use `read_csv` to import the data into a data frame called `df` with
default options except for the following:

i. The `IncidentDate` column should be recognized as containing calendar
dates via the option `col_date(format = "%d-%b-%y")` to the `cols(...)`
function, which is itself fed into the `col_types` option of `read_csv`.
This tells R that the incoming dates from the CSV file use the format
`"%d-%b-%y"`. (And remember that, once the data is imported correctly, R
will show dates with the `"%Y-%m-%d"` format.)

ii. The `IncidentTime` column should not be imported.

iii. The `Altitude` column should be read as `integer`. Note that
`integer` is not the same as `numeric` or `double`.

iv. The `Injury` column should be a factor column.

v. The following eight strings should be read as missing values:
`""`, `"NA"`, `"N/A"`, `"UNKN"`, `"Unknown"`, `"unknown"`, `"unkwn"`,
`"Unkwn"`

```{r}
library(readr)
my_col_types <-
    cols(
        IncidentDate = col_date(format = "%d-%b-%y"),
        IncidentTime = col_skip(),
        FlightID = col_character(),
        Aircraft = col_character(),
        Altitude = col_integer(),
        Airport = col_character(),
        LaserColor = col_character(),
        Injury = col_factor(),
        City = col_character(),
        State = col_character()
    )
my_na <- c("", "NA", "N/A", "UNKN", "Unknown", "unknown", "unkwn", "Unkwn")
df <- read_csv("lasers.csv", col_types = my_col_types, na = my_na)

```

### (b)

Save the number of rows in `df` in the variable `df_numrows`.

```{r}
df_numrows <- nrow(df)

```

### (c)

Save the total number of missing values in `df` in the variable
`df_missing`.

```{r}
df_missing <- sum(is.na(df))

```

### (d)

Ignoring `NA` values, calculate the median altitude (i.e., the median of
the `Altitude` column), and save it in the variable `df_medalt`.

```{r}
df_medalt <- median(df$Altitude, na.rm = TRUE)

```

### (e)

Plot a histogram of the natural logarithm (base e) of the `Altitude`
column. [Why the natural logarithm? It transforms the data so that the
histogram looks more like a normal distribution.]

```{r}
hist(log(df$Altitude))

```

### (f)

In a new data frame called `df_small`, save just the rows in positions
`1005`, `1010`, `1015`, ..., `1050` (by 5's in that order) and just
columns `FlightID` and `IncidentDate` (in that order).

```{r}
rows <- 1000 + 5 * (1:10)
df_small <- df[rows, c("FlightID", "IncidentDate")]

```

### (g)

Write out `df_small` in a CSV file called `lasers_small.csv`.

```{r}
write_csv(df_small, "lasers_small.csv")

```

## #6

The included file `ABBREV.txt` contains nutritional information from the
USDA. It is similar to a CSV file except that:

  * there are no column names (do not add any yourself)
  * the caret `^` is used instead of the comma `,`
  * the tilde `~` is used instead of the quotation mark `"`.

Taking into account these features, read and store the data in a
variable called `usda` using the `read_delim` function. Also make sure
that only the empty string is interpreted as a missing value. [Hint:
It may help to view the file first with a text editor such as Wordpad.
However, be sure not to overwrite the original file accidentally.]

```{r}
# Had to install tidyverse to get readr package
library(readr)
usda <- read_delim("ABBREV.txt", col_names = FALSE, delim = "^", quote = "~", na = "")

```

## #7

The included file `coffee.csv` contains coffee-consumption data by
country. The file is a regular CSV file with the following exceptions:
the delimiter is `;` and the decimal mark is `,` (instead of `.`).

### (a)

Load and save the data in a data frame called `coffee` using the
default options, except that you should include options to handle the
delimiter and decimal mark. [Hint: Use the option

      locale = locale(decimal_mark = ",")

to handle the decimal mark. In the console, `?locale` will provide
further information on the `locale()` command.]

```{r}
coffee <- read_delim("coffee.csv", delim = ";", locale = locale(decimal_mark = ","))

```

### (b)

Calculate the mean coffee consumption (ignoring missing values) and
store it in a variable called `mean_coffee`. [Note: The name of the
consumption column has a space in it. When dealing with spaces in column
names, R makes use of the "back tick" mark to make the full column name
clear.]

```{r}
mean_coffee <- mean(coffee$`Coffee Consumption`, na.rm = TRUE)

```
